\LARGE Performance Evaluation of Xenomai 3

### 2015-10-21 paper.tex
shorten the text

\begin{tabular}{|l| p{3cm} |l|}
\hline
Kernel & Patches & Source\\ \hline
mainline   & N/A & N/A \\ \hline
PREEMPT\_RT & patch-3.14.39-rt37 & \cite{p-rt} \\ \hline
Xenomai 2  & ipipe-core-3.14.39-arm-9 + xenomai v2.6.4 (Jul 13, 2015; HEAD) & \cite{p-ipipe} \cite{git-x2.6} \\ \hline
Xenomai 3  & ipipe-core-3.14.39-arm-9 + xenomai v3.0 (Oct 8, 2015) & \cite{p-ipipe} \cite{git-x3} \\ \hline
\end{tabular}\\
vspace{4mm} \\
Table 1: Different kernel configurations for the experiment\\

@@

\subsection{Experiment Result}
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_idle.jpg}
\caption{Cyclictest with no stress.}
\label{cyclictest-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_stress.jpg}
\caption{Cyclictest with stress.}
\label{cyclictest-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-idle.png}
\caption{Latency with no stress on Xenomai-2.}
\label{x2-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-cpu.png}
\caption{Latency with stress on Xenomai-2.}
\label{x2-latency-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-idle.png}
\caption{Latency with no stress on Xenomai-3.}
\label{x3-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-cpu.png}
\caption{Latency with stress on Xenomai-3.}
\label{x3-latency-stress}
\end{center}
\end{figure} 
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|r|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{idle}  \\ \cline{2-4} 
 & \multicolumn{1}{l|}{min} & \multicolumn{1}{l|}{avg} & \multicolumn{1}{l|}{max} \\ \hline
Mainline Linux & 39 & 43 & 1046  \\ \hline
Preempt-RT & 16 & 21 & 52 \\ \hline
Xenomai 2 & 8 & 9 & 41 \\ \hline
Xenomai 3 & 7 & 7 & 35 \\ \hline
 & \multicolumn{3}{c|}{cpu-stressed} \\ \hline
Mainline Linux & 39 & 52 & 1097 \\ \hline
Preempt-RT & 18 & 25 & 62 \\ \hline
Xenomai 2 & 8 & 19 & 45 \\ \hline
Xenomai 3 & 6 & 16 & 37 \\ \hline
\end{tabular}\\
vspace{4mm} \\
Table 2: cyclictest result on different kernel setups, in \textmu s\\
vspace{4mm} \\
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|c|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{user} \\ \cline{2-4} 
 & min & \multicolumn{1}{c|}{avg} & \multicolumn{1}{c|}{max} \\ \hline
Xenomai 2 & 8.541 & 9.458 & 34.583 \\ \hline
Xenomai 3 & 8.043 & 8.853 & 33.023 \\ \hline
& \multicolumn{3}{c|}{kernel}      \\ \hline
Xenomai 2 & 6.965 & 7.708 & 27.821 \\ \hline
Xenomai 3 & 5.567 & 6.479 & 25.577 \\ \hline
& \multicolumn{3}{c|}{timer-irq}   \\ \hline
Xenomai 2 & 2.129 & 2.654 & 11.343 \\ \hline
Xenomai 3 & 2.586 & 3.079 & 10.031 \\ \hline
vspace{4mm} \\
\end{tabular} \\
Table 3: latency result on different kernel setups, in \textmu s\\

From the data above, it shows that the latency performance of timer-irq in CPU-STRESSED situation is slightly worse on Xenomai-3(over 3 us) than on Xenomai-2(below 3 us). While this is not an occasional case(the same experiment was ran a couple of times), we tried to look into the test code of latency to figure out the cause of this phenomenon. In user-level mode, the measurement task is carried out by the function 'latency' inside the program; whereas in kernel or irq mode, the task is executed by issuing a Xenomai RTDM command such as 
\begin{verbatim}
rt_dev_ioctl(benchdev, \
RTTST_RTIOC_INTERM_BENCH_RES,&result); 
\end{verbatim} 
in an infinite loop to periodically launch timer and capture latencies. Unfortunately, we were unable to discover the reason for the discrepancy on the observed data.
\subsection{Experiment Result Analysis}
Our data shows that the average user-level latency in either Xenomai 2 or 3 is around 9 \textmu s in idle condition, 18 \textmu s in stressed condition. And in the worse case, the latency is below 50 \textmu s. Meanwhile, the latency of PREEMPT\_RT has a user-level latency more than 20 \textmu s, which means it can not guarantee a 50 \textmu s max latency (Actually in some cases the latency will exceed 100 \textmu s in stressed condition).  As for mainline Linux, the max latency easily exceeds 1 ms. When comparing the performance of Xenomai 3 to Xenomai 2, we can see that the statistics in cyclictest and latency test is slightly better, except for the case of timer-irq latency in stressed condition.

@@

The Linux kernel running under Xenomai 3 can be PREEMPT\_RT as well, which means Linux ROOT gets currently running thread's priority through the native option of Xenomai 3 for maximum API/ABI compatibility in real-time environments. For the tight latency requirements such as real-time networking, Xenomai 3 allows to switch to dual kernel configurations utilizing I-pipe for maximum real-time performance, which brings sligtly better performance than Xenomai 2. However, comparing to Xenomai 2, our results show the unexpected slowdown of timer-irq latency in stressed condition while Xenomai 3 was configured as the dual kernel configuration of Xenomai 3 though. From our performance evaluation, Xenomai 3 is known as the most flexible real-time option.

>>>>

\begin{tabular}{|l| p{3cm} |l|}
\hline
Kernel & Patches & Source\\ \hline
mainline   & N/A & N/A \\ \hline
PREEMPT\_RT & patch-3.14.39-rt37 & \cite{p-rt} \\ \hline
Xenomai 2  & ipipe-core-3.14.39-arm-9 + xenomai v2.6.4 (Jul 13, 2015; HEAD) & \cite{p-ipipe} \cite{git-x2.6} \\ \hline
Xenomai 3  & ipipe-core-3.14.39-arm-9 + xenomai v3.0 (Oct 8, 2015) & \cite{p-ipipe} \cite{git-x3} \\ \hline
\end{tabular}\\
\vspace{4mm} \\
Table 1: Different kernel configurations for the experiment\\

@@

\subsection{Experiment Result}
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_idle.jpg}
\caption{Cyclictest with no stress.}
\label{cyclictest-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_stress.jpg}
\caption{Cyclictest with stress.}
\label{cyclictest-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-idle.png}
\caption{Latency with no stress on Xenomai-2.}
\label{x2-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-cpu.png}
\caption{Latency with stress on Xenomai-2.}
\label{x2-latency-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-idle.png}
\caption{Latency with no stress on Xenomai-3.}
\label{x3-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-cpu.png}
\caption{Latency with stress on Xenomai-3.}
\label{x3-latency-stress}
\end{center}
\end{figure} 
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|r|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{idle}  \\ \cline{2-4} 
 & \multicolumn{1}{l|}{min} & \multicolumn{1}{l|}{avg} & \multicolumn{1}{l|}{max} \\ \hline
Mainline Linux & 39 & 43 & 1046  \\ \hline
Preempt-RT & 16 & 21 & 52 \\ \hline
Xenomai 2 & 8 & 9 & 41 \\ \hline
Xenomai 3 & 7 & 7 & 35 \\ \hline
 & \multicolumn{3}{c|}{cpu-stressed} \\ \hline
Mainline Linux & 39 & 52 & 1097 \\ \hline
Preempt-RT & 18 & 25 & 62 \\ \hline
Xenomai 2 & 8 & 19 & 45 \\ \hline
Xenomai 3 & 6 & 16 & 37 \\ \hline
\end{tabular}\\
\vspace{4mm} \\
Table 2: cyclictest result on different kernel setups, in \textmu s\\
\vspace{4mm} \\
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|c|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{user} \\ \cline{2-4} 
 & min & \multicolumn{1}{c|}{avg} & \multicolumn{1}{c|}{max} \\ \hline
Xenomai 2 & 8.541 & 9.458 & 34.583 \\ \hline
Xenomai 3 & 8.043 & 8.853 & 33.023 \\ \hline
& \multicolumn{3}{c|}{kernel}      \\ \hline
Xenomai 2 & 6.965 & 7.708 & 27.821 \\ \hline
Xenomai 3 & 5.567 & 6.479 & 25.577 \\ \hline
& \multicolumn{3}{c|}{timer-irq}   \\ \hline
Xenomai 2 & 2.129 & 2.654 & 11.343 \\ \hline
Xenomai 3 & 2.586 & 3.079 & 10.031 \\ \hline
vspace{4mm} \\
\end{tabular} \\
Table 3: latency result on different kernel setups, in \textmu s\\

From the data above, it shows that the latency performance of timer-irq in CPU-STRESSED situation is slightly worse on Xenomai-3(over 3 us) than on Xenomai-2(below 3 us). While this is not an occasional case(the same experiment was ran a couple of times), we tried to look into the test code of latency to figure out the cause of this phenomenon. In user-level mode, the measurement task is carried out by the function 'latency' inside the program; whereas in kernel or irq mode, the task is executed by issuing a Xenomai RTDM command such as 
\begin{verbatim}
rt_dev_ioctl(benchdev, \
RTTST_RTIOC_INTERM_BENCH_RES,&result); 
\end{verbatim} 
in an infinite loop to periodically launch timer and capture latencies. Unfortunately, we were unable to discover the reason for the discrepancy on the observed data.
\subsection{Experiment Result Analysis}
Our data shows that the average user-level latency in either Xenomai 2 or 3 is around 9 \textmu s in idle condition, 18 \textmu s in stressed condition. And in the worse case, the latency is below 50 \textmu s. Meanwhile, the latency of PREEMPT\_RT has an averaged user-level latency more than 20 \textmu s, which means it can not guarantee a 50 \textmu s max latency (Actually in some cases the latency will exceed 100 \textmu s in stressed condition).  As for mainline Linux, the max latency easily exceeds 1 ms. When comparing the performance of Xenomai 3 to Xenomai 2, we can see that the statistics in cyclictest and latency test is slightly better, except for the case of timer-irq latency in stressed condition.

@@

The Linux kernel running under Xenomai 3 can be PREEMPT\_RT as well through the native option of Xenomai 3 for maximum API/ABI compatibility in real-time environments. For the tight latency requirements such as real-time networking, Xenomai 3 allows to switch to dual kernel configurations utilizing I-pipe for maximum real-time performance, which brings sligtly better performance than Xenomai 2. However, comparing to Xenomai 2, our results show the unexpected slowdown of timer irq latency in stressed condition while Xenomai 3 was configured as the dual kernel configuration.
